---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip('activity.zip')
activity<-read.csv('activity.csv')
```
## What is mean total number of steps taken per day?
### sort out data
```{r}
library(dplyr)
stepbydate<-select(activity, date, steps)%>%
        group_by(date)%>%
        summarize(totalstep=sum(steps, na.rm = T))
```
### histogram and mean/median
```{r}
hist(stepbydate$totalstep,
     breaks = 100,
     xlab = "number of total steps per day",
     main = "histogram of the number of total steps per day");
meanstep<-mean(stepbydate$totalstep);
medianstep<-median(stepbydate$totalstep);
```
the mean total number of steps taken per day is `r round(meanstep,1)`, and the median total number of steps taken per day is `r round(medianstep,1)`

## What is the average daily activity pattern?
### pattern dataset
```{r}
pattern<-select(activity, interval, steps)%>%
        group_by(interval)%>%
        summarise(avgstep=mean(steps, na.rm = T))
```
### plotting and max interval
```{r}
plot(pattern$interval,pattern$avgstep,type = "l",
     xlab = "interval",
     ylab = "average steps",
     main="step patterns")
```
### find the interval with max avgsteps
```{r}
maxstep<-max(pattern$avgstep)
maxinterval<-pattern[pattern$avgstep==maxstep,]$interval
```
the interval with max average steps is `r maxinterval`.
 
## Imputing missing values
### summary of the data
```{r}
missingsteps<-sum(is.na(activity$steps));
missingdate<-sum(is.na(activity$date));
missinginterval<-sum(is.na(activity$interval));
```
### replace missing values in steps with mean steps and create a new dataset
```{r}
meansteptotal<-mean(activity$steps,na.rm = T)
activity_nareplace<-activity %>%
        mutate(nareplace=ifelse(is.na(steps),meansteptotal,steps))
```
### histogram
```{r}
stepbydatenarm<-select(activity_nareplace,nareplace,date)%>%
        group_by(date)%>%
        summarise(totalsteps=sum(nareplace));
 hist(stepbydatenarm$totalsteps,
      breaks = 100,
     xlab = "number of total steps per day",
     main = "histogram of the number of total steps per day");
```

### calculate mean and median
```{r}
meanstep_narm<-mean(stepbydatenarm$totalsteps)
medianstep_narm<-median(stepbydatenarm$totalsteps)
```
the mean total number of steps taken per day after na removal is `r round(meanstep_narm,0)`, the median total number of steps taken per day after na removal is `r round(medianstep_narm,0)`.
The values are changed, but did not biased from original data a lot.

## Are there differences in activity patterns between weekdays and weekends?
### create new variable to indicate weekdays or weekend
```{r}
activity_weekdays<-mutate(activity, weekdays=weekdays(as.Date(activity$date)),
                          weekdayorend=ifelse(weekdays=="Saturday"|weekdays=="Sunday","weekend","weekday"));
activity_weekdays<-select(activity_weekdays,interval,steps, weekdayorend)%>%
        group_by(interval)%>%
        mutate(avgsteps=mean(steps,na.rm = T));
```

### plotting
```{r}
library(ggplot2)
g<-ggplot(activity_weekdays, aes(x=interval, y=avgsteps), fill='transparent')
        g+geom_line(color='blue')+
        facet_wrap(.~weekdayorend,ncol = 1)+
        labs(y='number of steps')+
                theme(panel.background = element_rect(fill = "transparent",colour = NA))
          
```
